How do school travel planning stakeholders frame active school travel in Ontario, Canada?

### 2021-06-24 AST-Framing-Ontario/AST-Framing-Ontario.Rmd
Revised Introduction section

Factors that influence AST are typically conceptualized according to the socioecological model whereby children's travel behaviour is understood within the context of their household, social, and built environments [see @mitraIndependentMobilityMode2013]. At the individual level, older child age is often associated with AST [@mammenUnderstandingDriveEscort2012; @starkExploringIndependentActive2018; @wilsonUnderstandingChildParent2018]. There is some evidence that gender is a determinant of AST [@larsenInfluencePhysicalEnvironment2009], although this is not a strong or consistent finding [@rothmanDeclineActiveSchool2018; @schoeppeAssociationsChildrenActive2015]. Distance between home and school is associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school. Car ownership is an important household-level influence on AST [@pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], as is household income [@pontEnvironmentalCorrelatesChildren2009]. Parental perceptions of the environment [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010] and children's skills [@ghekiereInsightsChildrenIndependent2017] also influence whether they allow their children to walk or cycle to school. Finally, many studies have found that the quality of the built environment [@ikedaAssociationsChildrenActive2018; @kimBuiltNaturalEnvironmental2020] and active travel infrastructure [@pontEnvironmentalCorrelatesChildren2009] facilitate AST. Concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012]. The consensus is that all of these factors are interrelated and that interventions ought to target multiple factors in order to increase walking and bicycling levels to school.

@@

AST is one of the most common opportunities for children to travel without adult supervision on a regular basis. Children's independent mobility (CIM) was first described by Hillman et al. [-@hillmanOneFalseMove1990] as the freedom of children to travel and play within their neighborhood and city without the presence of adults. The ability of children to reach destinations such as school or local parks by walking or bicycling on their own as a growing area of research that is related to the body of literature on AST. Similar to AST, CIM is much less common now than it was for previous generations [@bhosaleIntergenerationalChangeChildren2017], and is influenced by many similar personal and environmental factors as AST [@marziChildrenIndependentMobility2018; @mitraIndependentMobilityMode2013]. In a study conducted in multiple Canadian cities, Riazi et al. [-@riaziCorrelatesChildrenIndependent2019] found that a range of factors such as child grade, car ownership, and parental perceptions of safety and environment were associated with CIM.

@@

The desire to increase AST is warranted - there is strong evidence that children who walk and bicycle to school accrue physical and mental health benefits. Many studies and reviews have focused on the association between AST or CIM and physical activity [e.g., @faulknerActiveSchoolTransport2009; @mitraChildrenActivitytransportationLifestyles2017; @schoeppeAssociationsChildrenActive2015], with findings consistently demonstrating that children who travel by walking or bicycling to school are more active than their peers who do not use active travel. More recently, researchers have been exploring the link between transport and children's wellbeing [@waygoodTransportChildrenWellbeing2020], which has relevant applications to the study of school travel and satisfaction [see @vandenbergFactorsAffectingParental2020]. CIM is also important for different domains of children's health and wellbeing [@riaziChildrenIndependentMobility2018]. It offers benefits such as increasing traffic and safety skills, boosting spatial awareness when navigating public spaces, and providing more opportunity for social interaction with peers [@schoeppeAssociationsChildrenIndependent2014].

@@

In the case of AST and CIM in Canada, Riazi et al. [-@riaziCorrelatesChildrenIndependent2019] state: "it will be vital for interventions to target modifiable factors, including children's and parents' perceptions of their social environment." Parents play the important role of "gatekeeper" by either granting or restricting CIM licenses, meaning whether children can travel alone [@sharminAssociationBuiltEnvironment2017]. For this reason, stakeholders involved in STP produce information targeted for both children and parents, and aim to involve the broader community in their efforts.

@@

Any goals for AST, and also CIM, ought to be clearly articulated and reflected in transportation plans and policies to guide initiatives. Presently, however, there has been no study to date that explores how Canadian municipalities and schools frame and discuss AST with the public. Framing or content analysis is one method to analyze how particular issues are framed to groups of people, for instance parents or educators who might be inclined to support AST. It attempts to understand how information presented from a "communicator" leads the "receiver" to a desired response [@reynardGrowthResilienceHow2021]. In a recent paper [@reynardGrowthResilienceHow2021], framing analysis was applied to review municipal policies addressing climate change in four western Canadian cities. The authors determined that most issues are framed favouring merely a tinkering of the status quo as opposed to pushing for real change. Natural language processing (NLP) was used for a similar purpose to examine content in general plans from Californian cities [@brinkleyWhatPlanUsing2021]. The way policy issues are framed is important to understand because it plays a role in either altering or preserving the existing social perceptions.

>>>>

Factors that influence AST are typically conceptualized according to the socioecological model whereby children's travel behaviour is understood within the context of their household, social, and built environments [see @mitraIndependentMobilityMode2013]. At the individual level, older child age is often associated with AST [@mammenUnderstandingDriveEscort2012; @starkExploringIndependentActive2018; @wilsonUnderstandingChildParent2018]. There is some evidence that gender is a determinant of AST [@babbChildrenActiveTravel2017; @larsenInfluencePhysicalEnvironment2009], although this is not a strong or consistent finding [@rothmanDeclineActiveSchool2018; @schoeppeAssociationsChildrenActive2015]. Distance between home and school is associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school. Car ownership is an important household-level influence on AST [@pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], as is household income [@pontEnvironmentalCorrelatesChildren2009]. Parental perceptions of the environment [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010] and children's skills [@ghekiereInsightsChildrenIndependent2017] also influence whether they allow their children to walk or cycle to school. Finally, many studies have found that the quality of the built environment [@ikedaAssociationsChildrenActive2018; @kimBuiltNaturalEnvironmental2020] and active travel infrastructure [@pontEnvironmentalCorrelatesChildren2009] facilitate AST. Concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012]. The consensus is that all of these factors are interrelated and that interventions ought to target multiple factors in order to increase walking and bicycling levels to school.

@@

AST is one of the most common opportunities for children to travel without adult supervision on a regular basis. Children's independent mobility (CIM) was first described by Hillman et al. [-@hillmanOneFalseMove1990] as the freedom of children to travel and play within their neighborhood and city without the presence of adults. The ability of children to reach destinations such as school or local parks by walking or bicycling on their own as a growing area of research that is related to the body of literature on AST. Similar to AST, CIM is much less common now than it was for previous generations [@bhosaleIntergenerationalChangeChildren2017], and is influenced by many similar personal and environmental factors as AST [@marziChildrenIndependentMobility2018; @mitraIndependentMobilityMode2013]. In a study conducted in multiple Canadian cities, Riazi et al. [-@riaziCorrelatesChildrenIndependent2019] found that a range of factors such as child grade, car ownership, and parental perceptions of safety and environment were associated with CIM. In addition to independent trips to school, children also travel by walking or bicycling on their own to other non-school destinations such as outdoor spaces (i.e., parks, playgrounds), the homes of family members or friends, shopping places, and libraries, to name a few [@badlandAssessingNeighbourhoodDestination2015; @villanuevaWhereChildrenTravel2013a].

@@

The desire to increase AST is warranted - there is strong evidence that children who walk and bicycle to school accrue physical and mental health benefits. Many studies and reviews have focused on the association between AST or CIM and physical activity [e.g., @faulknerActiveSchoolTransport2009; @mitraChildrenActivitytransportationLifestyles2017; @schoeppeAssociationsChildrenActive2015], with findings consistently demonstrating that children who travel by walking or bicycling to school are more active than their peers who do not use active travel. More recently, researchers have been exploring the link between transport and children's wellbeing [@babbChildrenActiveTravel2017; @waygoodTransportChildrenWellbeing2020], which has relevant applications to the study of school travel and satisfaction [see @vandenbergFactorsAffectingParental2020]. CIM is also important for different domains of children's health and wellbeing [@riaziChildrenIndependentMobility2018]. It offers benefits such as increasing traffic and safety skills, boosting spatial awareness when navigating public spaces, and providing more opportunity for social interaction with peers [@schoeppeAssociationsChildrenIndependent2014].

@@

In the case of AST and CIM in Canada, Riazi et al. [-@riaziCorrelatesChildrenIndependent2019] state: "it will be vital for interventions to target modifiable factors, including children's and parents' perceptions of their social environment." Parents play the important role of "gatekeeper" by either granting or restricting CIM licenses, meaning whether children can travel alone [@sharminAssociationBuiltEnvironment2017]. For this reason, stakeholders involved in STP produce information targeted for both children and parents, and aim to involve the broader community in their efforts to increase the number of children walking and bicycling to school..

@@

Any goals for AST, and also CIM, ought to be clearly articulated and reflected in transportation plans and policies to guide initiatives. Presently, however, there has been no study to date that explores how Canadian municipalities and schools frame and discuss AST with the public. Content analysis is one method to analyze how particular issues are framed to groups of people, for instance parents or educators who might be inclined to support AST. It attempts to understand how information presented from a "communicator" leads the "receiver" to a desired response [@reynardGrowthResilienceHow2021]. In a recent paper [@reynardGrowthResilienceHow2021], framing analysis was applied to review municipal policies addressing climate change in four western Canadian cities. Natural language processing (NLP) was used for a similar purpose to examine content in general plans from Californian cities [@brinkleyWhatPlanUsing2021]. The way policy issues are framed is important to understand because it plays a role in either altering or preserving the existing social perceptions.
